### 2019-05-16 README.Rmd
Initial commit

---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
# Please put your title here to include it in the file below. 
Title <- "Canary in the Coal Mine: Anti-NGO Legislation as a Predictor of Civil Society Repression"
Authors <- "Suparna Chaudhry and Andrew Heiss"
Year <- "2019"
```

# `r Title`

[Suparna Chaudhry](http://www.suparnachaudhry.com/) • Department of Political Science • Christopher Newport University  
[Andrew Heiss](https://www.andrewheiss.com/) • Andrew Young School of Policy Studies • Georgia State University

# Abstract

Many countries across the globe are increasingly cracking down on civil society groups through legal restrictions. While research on democratic backsliding has tended to focus on rights and institutions in the electoral arena, this new wave of laws—many of which appear mundane on their face—represent a bureaucratic form of repressive that could be indicative of more severe human rights abuses and backsliding. This is especially the case for electoral democracies, which unlike autocracies do not aggressively and openly attack civic space. In this paper, we ask: do legal crackdowns on NGOs predict broader civil society repression? We argue that this is the case because anti-NGO laws are among the most subtle means of repression and attract the least least domestic and international condemnation, as many audiences mistake this repression as regulation and are not personally affected by it. These laws also make it easier for states to prevent and deter future large-scale popular mobilization. Using original data on civil society restrictions over the last three decades, we test whether NGO crackdown is a predictor of attacks on journalists, independent media, academic freedom, and freedoms of expression and association. This project has important implications: if crackdown on NGOs is indeed an early warning, it can potentially prevent further deteriorations in space for civil society.

---

This repository contains the data and code for our paper:

> `r Authors`. (`r Year`). *`r Title`*. Accessed `r format(Sys.time(), '%B %e, %Y')`.

Our pre-print is online here:

> Authors, (YYYY). _`r Title`_. Name of journal/book, Accessed `r format(Sys.Date(), "%d %b %Y")`. Online at <https://doi.org/xxx/xxx>

## How to cite

Please cite this compendium as:

> `r Authors`. (`r format(Sys.Date(), "%Y")`). *Compendium of R code and data for `r Title`*. Accessed `r format(Sys.time(), '%B %e, %Y')`.

### How to download or install

You can either [download the compendium as a ZIP file](/archive/master.zip) or use GitHub to clone or fork the compendium repository (see the green "Clone or download" button at the top of the GitHub page).

In order to reproduce this project, you'll need to install the compendium as an R package. After downloading the compendium, do the following:

### Licenses

**Text and figures:** All prose and images are licensed under Creative Commons ([CC-BY-4.0](http://creativecommons.org/licenses/by/4.0/))

**Code:** All code is licensed under the [MIT License](LICENSE.md).

### Contributions

We welcome contributions from everyone. Before you get started, please see our [contributor guidelines](CONTRIBUTING.md). Please note that this project is released with a [Contributor Code of Conduct](CONDUCT.md). By participating in this project you agree to abide by its terms.