### 2020-03-08 analysis/03_models.Rmd
Add TODOs for models and predictions

---
title: "Models"
author: "Suparna Chaudhry and Andrew Heiss"
date: "Last run: `r format(Sys.time(), '%F')`"
---

```{r setup, warning=FALSE, message=FALSE}
library(tidyverse)
library(here)

canary <- read_rds(here("data", "derived_data", "canary_clean.rds"))
```

TODO:

- Causal model
- Run actual models
- Show marginal effects - change in predicted probability of PTS levels as CS restriction changes
- Use matching (https://cran.r-project.org/web/packages/PanelMatch/vignettes/using_panelmatch.html), IPW (BlackwellGlynn:2018) to isolate effect
- Careful selection of confounders, following BellemareMasakiPepinsky:2017 and Samii:2016
- Use MSMs or whatever BlackwellGlynn:2018 suggest?