### 2020-03-08 analysis/04_predictions.Rmd

---
title: "Predictions"
author: "Suparna Chaudhry and Andrew Heiss"
date: "Last run: `r format(Sys.time(), '%F')`"
---

```{r setup, warning=FALSE, message=FALSE}
library(tidyverse)
library(here)

canary <- read_rds(here("data", "derived_data", "canary_clean.rds"))
```

TODO:

- Split into training/test data
- Generate predictions
- Compare correctness, AUC, etc. for models with and without CS restrictions