### 2021-11-12 manuscript/manuscript.Rmd
Address R2's issues 10, 11, 12, and 14—RIP V-Dem repression :(

---
title: "NGO Repression as a Predictor of Worsening Human Rights Abuses"
short-title: "NGO Repression and Human Rights"
author:
- name: Suparna Chaudhry
  affiliation: Lewis & Clark College
  email: schaudhry@lclark.edu
  url: http://www.suparnachaudhry.com/
- name: Andrew Heiss
  affiliation: Georgia State University
  email: aheiss@gsu.edu
  url: https://www.andrewheiss.com/
date: "July 20, 2021"
published: Under review
code-repo: "Access the code and full analysis notebook at <https://stats.andrewheiss.com/canary-ngos/>"
abstract: >-
  An increasing number of countries have recently cracked down on non-governmental organizations (NGOs). Much of this crackdown is sanctioned by law and represents a bureaucratic form of repression that could indicate more severe human rights abuses in the future. This is especially the case for democracies, which unlike autocracies, may not aggressively attack civic space. We explore if crackdowns on NGOs predict broader human rights repression. Anti-NGO laws are among the most subtle means of repression and attract lesser domestic and international condemnation compared to the use of violence. Using original data on NGO repression, we test whether NGO crackdown is a predictor of political terror, and violations of physical integrity rights and civil liberties. We find that while formal de jure laws provide little information in predicting future repression, their patterns of implementation, or de facto civil society repression, predicts worsening respect for physical integrity rights and civil liberties.
# keywords: civil society, NGOs, repression, human rights, political terror, civil liberties
word-count: "7,159"
thanks: >-
  We would like to thank participants of the 2021 International Studies Association (ISA) convention, and the "Human Rights on the Edge: The Future of International Human Rights Law and Practice" workshop hosted by the Global Human Rights Hub at Arizona State University for their comments and feedback.
reference-section-title: References
bibliography: bibliography.bib
toc: false
title-page: true
endnotes: true
endfloat: false
mainfont: Cochineal
sansfont: Inter
author-note: >-
  Suparna Chaudhry, PhD, Assistant Professor of International Affairs, Department of International Affairs, Lewis \& Clark College; Andrew Heiss, PhD, Assistant Professor of Public Management, Andrew Young School of Policy Studies, Georgia State University. 
correspondence: >-
  Correspondence concerning this article should be addressed to Andrew Heiss, PhD, Department of Public Management and Policy, Andrew Young School of Policy Studies, Georgia State University, 14 Marietta Street NW, Suite 357, Atlanta, GA 30303. Contact: aheiss@gsu.edu
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
# Figure out output format
is_docx <- knitr::pandoc_to("docx") | knitr::pandoc_to("odt")
is_latex <- knitr::pandoc_to("latex")
is_html <- knitr::pandoc_to("html")

# Word-specific things
table_format <- ifelse(is_docx, "huxtable", "kableExtra")  # Huxtable tables
conditional_dpi <- ifelse(is_docx, 300, 300)  # Higher DPI
conditional_align <- ifelse(is_docx, "default", "center")  # Word doesn't support align

# Knitr options
knitr::opts_chunk$set(
  echo = FALSE, warning = FALSE, message = FALSE,
  tidy.opts = list(width.cutoff = 120),  # Code width
  fig.retina = 3, dpi = conditional_dpi,
  fig.width = 7, fig.asp = 0.618,
  fig.align = conditional_align, out.width = "100%",
  fig.path = "output/figures/",
  cache.path = "output/_cache/",
  fig.process = function(x) {  # Remove "-1" from figure names
    x2 = sub('-\\d+([.][a-z]+)$', '\\1', x)
    if (file.rename(x, x2)) x2 else x
  }
)

# R options
options(
  width = 90,  # Output width
  dplyr.summarise.inform = FALSE,  # Turn off dplyr's summarize() auto messages
  knitr.kable.NA = "",  # Make NAs blank in kables
  kableExtra.latex.load_packages = FALSE,  # Don't add LaTeX preamble stuff
  modelsummary_factory_default = table_format,  # Set modelsummary backend
  modelsummary_format_numeric_latex = "plain"  # Don't use siunitx
)

# modelsummary regression table notes
if (is_latex) {
  model_note <- "Posterior means; 95\\\\% credible intervals in brackets" 
} else {
  model_note <- "Posterior means; 95% credible intervals in brackets"
}
```

```{r libraries-data, include=FALSE}
library(tidyverse)
library(targets)
library(kableExtra)
library(modelsummary)
library(patchwork)
library(scales)
library(ggtext)
library(sf)
library(here)
library(brms)
library(emmeans)
library(broom.mixed)

withr::with_dir(here::here(), {
  source(tar_read(plot_funs))

  canary <- tar_read(panel)
  tar_load(civicus_map_data)
  tar_load(civicus_clean)

  tar_load(c(m_pts_total, m_pts_advocacy, m_pts_entry, m_pts_funding,
             m_clphy_total, m_clphy_advocacy, m_clphy_entry, m_clphy_funding,
             m_clpriv_total, m_clpriv_advocacy, m_clpriv_entry, m_clpriv_funding,
             m_pts_v2csreprss, m_clphy_v2csreprss, m_clpriv_v2csreprss, m_lhr_v2csreprss))

  # Model tables
  tar_load(c(models_paper_pts, models_paper_clphy, 
             models_paper_clpriv, models_paper_lhr))

  # Load lookup list for coefficients in regression tables
  tar_load(coef_list)

  # Load marginal effects
  tar_load(c(mfx_e1a_pts, mfx_e1b_clphy, mfx_e1c_clpriv, mfx_e1d_lhr,
             mfx_e2a_pts, mfx_e2b_clphy, mfx_e2c_clpriv, mfx_e2d_lhr))
})
```

```{r calc-civicus}
civicus <- civicus_clean %>%
  count(category) %>% 
  mutate(prop = n / sum(n)) %>% 
  mutate(prop_nice = percent_format(accuracy = 1)(prop)) %>% 
  split(.$category)
```

\memSingle

Over the last few decades, governments across the globe have been increasingly cracking down on civil society organizations (CSOs), particularly non-governmental organizations (NGOs)—a phenomenon commonly known as closing or shrinking civic space [@CarothersBrechenmacher:2014; @DPR:2016; @Chaudhry:2016; @DupuyPrakash:2018]. More than 100 countries have proposed or enacted 244 different measures that have restricted, repressed, or shut down civil society since 2013 [@icnl:2021]. In Human Rights Watch's 2016 World Report, Executive Director Kenneth Roth argued that civil society was under more aggressive attack than at any time in recent memory, and over the past decade, the majority of the world has seen a substantial narrowing of allowable civic space. As seen in Figure \@ref(fig:civicus-map), by the end of 2020, only `r civicus$Open$prop_nice` of countries had open and unrestricted civic space [@civicusmonitor; @ChaudhryHeiss:2021].[^open]

[^open]: Following CIVICUS, open civic space in a country indicates that the state enables and safeguards citizens' freedom to form associations, authorities are tolerant of criticism from CSOs, and that laws governing freedom of peaceful assembly adhere to international laws and standards.

```{r civicus-map, fig.cap="2020 CIVICUS Monitor civic space ratings", fig.pos="h"}
ggplot() +
  geom_sf(data = civicus_map_data, aes(fill = fct_rev(category)), size = 0.15, color = "black") +
  coord_sf(crs = st_crs("ESRI:54030"), datum = NA) +  # Robinson
  scale_fill_viridis_d(option = "rocket", direction = -1,
                       na.translate = FALSE, name = "Civic space") +
  theme_void(base_size = 9, base_family = "Inter Light") +
  theme(legend.position = "bottom",
        legend.key.size = unit(0.7, "lines"))
```

This hostility to civil society is not unique to autocratic governments or even illiberal democracies [@AbramowitzSchenkkan:2018]—long-established liberal democracies have also increased restrictions on civil society in their countries. Over the past decade, India has canceled the licenses of tens of thousands of NGOs, targeted major groups such as Greenpeace International, Amnesty International and Ford Foundation, and has banned thousands of NGOs from receiving foreign funds [@press-trust-india:2015; @press-trust-india:2019]. In 2013 Canada attempted to revoke the charitable status of NGOs for opposing the Northern Gateway pipeline and Canadian mining companies' initiatives abroad [@CIVICUS:2013]. In 2019, activists from the US-based nonprofit No More Deaths were arrested for providing food and water to migrants at the US-Mexico border and faced a 20-year prison sentence before their convictions were overturned in 2020 [@Cramer:2020].

Much of this targeting of NGOs has been non-violent. Rather than publicly arrest or beat activists, states implement "administrative crackdown," or the passage of laws to create barriers to NGO advocacy, entry, and funding [@Chaudhry:2016]. The use of legal tools instead of violent repression means that NGOs have to adapt their practices to stay within the bounds of permitted activities and operations [@Heiss:2019; @Heiss:2017]. While this can lead to the depoliticization and avoidance of direct political activism by CSOs [@BloodgoodTremblay-Boire:2017], states cannot soften or neutralize all forms of political activity and advocacy in subtle, legal ways like they have with NGOs. Sooner or later, states may need to deal with threats to their stability in a more direct, violent, confrontational manner. Anti-NGO laws could be an important early warning sign—or a canary in the coal mine—of worsening human rights repression, signaling that a state will continue down the path of further repression. What does state targeting of NGOs tell us about the state's broader respect for civil society and the overall level of human rights repression? Can the crackdown on NGOs act as a predictor of subsequent increasing repression?

In this paper, we explore the power of anti-NGO repression in predicting political terror and underlying human rights abuses. We argue that information about NGO crackdown can be a valuable indicator of human rights trajectories. It can act as an early warning signal for worsening repression, suggesting which countries are testing the waters with comparatively "milder" methods of crackdown before enacting more violent forms of repression. To assess if attacks on NGOs and activists can provide us useful information about future repression, we use an original dataset on state crackdown on NGOs using legal repression across all countries from 1990–2013. We differentiate between two forms of NGO repression: official de jure anti-NGO restrictions that are codified in law, and their de facto implementation. Using multilevel Bayesian modeling, we find that formal de jure laws potentially provide a weak signal of future repression. However, we find that accounting for their implementation does boost our models' predictive power. Increasing de facto civil society repression, on the other hand, predicts both a moderate increase in the probability of moderate political terror and a deterioration in the overall respect for human rights in a typical country.

Our paper makes an important contribution in predicting the onset of repressive state policies. While structural factors such as levels of democracy, economic development, electoral competitiveness can identify countries that are likely to have more or less respect for civic freedoms, these factors are also slow to change over time. As such, they may not be particularly suitable for forecasting short-term trajectories, especially for moderately repressive countries. On the other hand, state repression of NGOs is a more dynamic process that may provide insight into changing levels of human rights standards and repression in subsequent years. While recent research has made important contributions to studying the extent and effects of state crackdown on NGOs, much of this focuses on de jure restrictions and its impact on NGOs and donors [@christensen2013; @DPR:2016; @DupuyPrakash:2018; @BromleySchoferLonghofer:2020; @BakkeMitchellSmidt:2020; @SmidtPereraMitchell:2021]. By looking at laws *and* their implementation, as well as the extra-legal treatment of NGOs, we are able to assess the long-term effects of closing civic space on broader patterns of repression.

Below, we first look at existing literature on repression and political violence, outlining state motivations to repress NGOs. We then elaborate on why attacking NGOs may be an important predictor of subsequent repression. Using both original data as well as data from several existing datasets, we build an ensemble of predictive models and explore (1) the marginal effects of civil society repression on different forms of broader human rights repression, and (2) the out-of-sample predictive performance of each model. We conclude by discussing the implications of our findings on autocratic survival, and how states may manage NGOs as a strategy to remain in power.

# Government repression and NGO restrictions

Repression is the use of coercive action against an individual or organization, within the territorial jurisdiction of the state, for the purpose of imposing a cost on the target as well as deterring specific activities perceived to be challenging to government personnel, practices, or institutions [@Goldstein:1978, p. xxvii; @Davenport:2007, 487; and @ConradRitter:2019, 15]. Early literature on repression focused on the role of structural factors such as levels of democracy, economic development, and electoral competitiveness in predisposing a state to the increased use of repression. Most notably, democratic forms of government—and democratization in general—has a pacifying effect on state repression [@DavenportArmstrong:2004; @Krain:1997; @Richards:1999]. Democracies have a negative effect on repression largely because of the increased probability of sanctions against authorities for undesirable behavior [@Davenport:2007a, 47].

More recent research focuses on how the behavior of various actors shapes government response. One of the most consistent findings in the political violence literature is that authorities repress to counter ongoing challenges to the status quo. In other words, dissent increases repressive behavior [@Davenport:2007]. However, dissent and repression have an endogenous relationship: state authorities also frequently engage in preventive repression to undermine or restrict groups' abilities to dissent [@RitterConrad:2016; @ConradRitter:2019]. Such preventive repression also damages organizations' capacity to impose costs [@Davenport:2007a, 47]. NGOs are one actor that can pose challenges to political authorities, and state repression of these actors, especially through legal restrictions can be considered a form of preventive repression [@Chaudhry:2016].

This preventive repression of NGOs is a recent phenomenon. While many saw the initial proliferation of NGOs in the 1980s and 1990s as leading to a "symbiotic relationship" between NGOs and states due to their mutual goals, this perception changed over time [@Reimann:2006, 63]. With expanded powers and their perceived role in the Color Revolutions, NGOs began to be seen as having a real influence on domestic and international politics [@CarothersBrechenmacher:2014]. State perception of NGOS as threatening became pervasive as these organizations demonstrated their ability to influence electoral politics [@DPR:2016; @Weiss:2009], mobilize aid [@MurdieBhasin:2011; @BBCM2014; @Boulding:2014], and threaten a state's economic interests [@DietrichMurdie:2017; @LebovicVoeten:2009; @Nielsen:2013].

Given this threat, a number of states decided to repress threatening NGOs. While violence is often effective at curbing dissent, it can also backfire, leading to widespread protests, decreasing leaders' legitimacy and increasing their criminal liability [@KimSikkink:2010; @Sikkink:2011]. In contrast to violence, administrative crackdown has fewer adverse consequences—citizens may see it as regulation rather than repression, resulting in little to no counter-mobilization. It may also circumvent concerns among state security or bureaucratic agents about illegitimate orders based on violent repression [@StephanChenoweth:2008]. There are also fewer international consequences since laws—unlike violence—rarely invite condemnation or threats of aid withdrawal [@CarothersBrechenmacher:2014].

This administrative crackdown via anti-NGO laws may foreshadow future government transgressions. Autocratic leaders want to avoid accountability, and NGOs often stand in conflict with that goal. Civil society organizations, including NGOs, can help shape the narrative of leaders' performance both domestically and internationally. Monitoring by NGOs and subsequent shaming can include significant external revenue loss to states in the form of loss of aid, trade, and foreign direct investments [@KeckSikkink:1998; @BarryClayFlynn:2013]. Government attempts to repress NGOs can signal the beginning of a deteriorating human rights situation. Such repression reflects the government's willingness and capability to use measures that can make it avoid accountability. Once a state targets NGOs, over time it may grow less hesitant to apply violent repression more widely and more severely.

In exploring the impact of state crackdown on NGOs on subsequent human rights conditions, we look at both *de jure* measures—or formal legislation that seeks to repress NGOs—as well as the *de facto* implementation of these measures. This distinction is important—while states may pass anti-NGO laws, these laws can also lie dormant until governments feel the need to enforce them. For instance, Russia passed its 2015 Undesirable Organizations Law with the purpose of targeting and eliminating specific foreign-connected NGOs it deemed dangerous. One of the law's sponsors, Aleksandr Tarnavsky, described the law as a preventive measure that would not affect the majority of NGOs working in Russia. Rather, the law would be a "weapon hanging on the wall that never fires" and stand as a warning to potentially uncooperative NGOs [@Kozenko:2015]. Similarly, Egypt's 2002 NGO law instated severe barriers to freedom of association, banned NGOs from engaging in activities that "threaten national unity or violate public order or morality," and also prevented groups from receiving foreign funding. But despite the harsh legal framework, the Egyptian NGO sector nearly doubled in size between 1993–2011. This is because the state targeted a limited number of organizations working on politically sensitive issues such as human rights and democracy, while simultaneously tolerating organizations working on tamer issues [@Brechenmacher:2017].

Simply noting the passage of (or counting) these laws may therefore miss the effect of their de facto implementation. Moreover, the implementation of these laws may not be limited to its legal character. The government may also engage in harassment of NGOs and activists to dissuade NGOs from carrying out their operations. This may include measures such as detentions, short-term incarceration, as well as beatings, threats to families and destruction of valuable property.

Finally, counting official laws can also undermeasure civil society repression because governments can repress NGOs without relying on formal laws. As prior research in Egypt, Ethiopia, and Russia points out, "rather than systematically enforcing all repressive legal measures-often impossible owing to capacity constraints," governments may choose to use extra-legal means to intimidate, harass or prosecute a few NGOs or activists. These cases, especially during moments or crises or contention, "highlight to the wider NGO community-and citizenry-that the government is willing and capable of using the law to repress dissent" [@Brechenmacher:2017, 94].

Based on these various purposes of NGO restrictions and the different methods states use to implement them, we propose two empirical expectations about the relationship between civil society repression and general human rights:

- **Empirical expectation 1** (effect of de jure NGO repression): If a country imposes a formal anti-NGO law, then it is likely to signal future deteriorating human rights conditions.
- **Empirical expectation 2** (effect of de facto NGO repression): If a country implements its formal anti-NGO laws and represses civil society, then it is likely to signal future deteriorating human rights conditions.

# Data and modeling approach

To test these empirical expectations, we use a new dataset of civil society laws, as well as measures from several other larger datasets focused on human rights, political terror, and democratization. Table \@ref(tab:var-summary) provides a summary of our key variables.

(ref:pts) @GibneyCornettWood:2019

(ref:vdem) @vdem-v10

(ref:lhr) @FarissKenwickReuning:2020

(ref:chaudhry) @Chaudhry:2016

```{r var-summary}
# For whatever reason, bookdown::pdf_document2 can't process in-table citations
# like @vdem-v10; it just skips them. bookdown::html_document2 works just fine,
# though. This is a known issue (see
# https://github.com/haozhu233/kableExtra/issues/214#issuecomment-421706528),
# and the solution is to use bookdown text references
# (https://bookdown.org/yihui/bookdown/markdown-extensions-by-bookdown.html#text-references)
# out in the main text and then include the reference key in the table:
# 
# In the text:
# (ref:vdem) @vdem-v10
#
# In the table:
# (ref:vdem)
# 
# It's weird, but it works
#
# BUT it doesn't work with Word, and flextable specifically. (ref:vdem) gets
# correctly replaced with @vdem-10, but then it doesn't get processed by pandoc.
# One way around this is to use the ftExtra package, which has colformat_md()
# that will treat a column like Markdown and will allow pandoc to process
# citations (see https://github.com/davidgohel/flextable/issues/243 and
# ftExtra's citation support here:
# https://ftextra.atusy.net/articles/format_columns.html#citations). BUT BUT the
# ordering of bookdown's text references replacement messes with flextable's
# citation processing or something, so pandoc references that are inserted with
# (ref:vdem) remain as @vdem-10 when knitted. The way around *this* is to use
# the raw @citekey in the table. Oof.
#
# Here's a minimal example:
#
# bloop <- tribble(
#   ~Thing, ~Source, ~Source_word,
#   "This *is* a test", "(ref:vdem)", "@vdem-v10"
# ) 
# 
# if (is_docx) {
#   bloop %>% 
#     select(Thing, Source = Source_word) %>% 
#     hux() %>% 
#     huxtable::as_flextable() %>% 
#     ftExtra::colformat_md() 
# } else {
#   bloop %>% 
#     select(-Source_word) %>% 
#     kbl(booktabs = TRUE) %>% 
#     kable_styling()
# }

tbl_var_summary_caption <- "Outcomes and predictors"

tbl_var_summary <- tribble(
  ~Measure, ~Category, ~Description, ~Source,
  "Political Terror Scale (PTS)", NA, "Ordered category ranging from 1 (least terror) to 5 (most terror)", "(ref:pts)",
  "Latent human rights values", NA, "≈ −3–5; higher values indicate greater respect for human rights", "(ref:lhr)",
  "Total legal barriers", "De jure laws", "Count of all anti-NGO legal barriers", "(ref:chaudhry)",
  "Barriers to advocacy", "De jure laws", "Count of laws restricting NGO advocacy", "(ref:chaudhry)",
  "Barriers to entry", "De jure laws", "Count of laws restricting NGO entry, including registration", "(ref:chaudhry)",
  "Barriers to funding", "De jure laws", "Count of laws restricting NGO funding sources", "(ref:chaudhry)",
  "CSO repression", "De facto repression", "≈ −3–3; higher values indicate less repression", "(ref:vdem); v2csreprss"
)

if (!is_docx) {
  tbl_var_summary %>%
    kbl(caption = tbl_var_summary_caption,
        booktabs = TRUE, linesep = "\\addlinespace") %>% 
    pack_rows(index = c("Outcomes" = 2, "Predictors" = 5),
              indent = FALSE) %>% 
    kable_styling(font_size = ifelse(is_latex, 8.5, NA)) %>% 
    column_spec(1, width = "0.16\\\\textwidth") %>% 
    column_spec(2, width = "0.1\\\\textwidth") %>% 
    column_spec(3, width = "0.4\\\\textwidth") %>% 
    column_spec(4, width = "0.2\\\\textwidth")
} else {
  tbl_var_summary %>% 
    hux() %>% 
    set_font_size(10) %>%
    set_all_padding(3) %>% 
    insert_row("Outcomes", "", "", "", after = 1) %>% 
    insert_row("Predictors", "", "", "", after = 4) %>% 
    set_bold(1, everywhere) %>%
    set_bold(c(2, 5), everywhere) %>%
    set_bottom_border(c(2, 5), everywhere) %>%
    set_width(1) %>% 
    set_wrap(TRUE) %>% 
    huxtable::set_caption(tbl_var_summary_caption) %>% 
    huxtable::as_flextable()
}
```

## Measuring de jure and de facto civil society repression

To measure de jure repression, we use data on anti-NGO laws from @Chaudhry:2016 which includes official NGO laws for all countries (excluding micro-states) between 1990 and 2013. This measure of formal legal barriers does not necessarily represent individual laws. Instead, we count the presence or absence of specific types of anti-NGO legal barriers as a binary variable. If a country passes a broader NGO laws that implements several of these barriers simultaneously (e.g. making registration burdensome and restricting foreign funds), the values for each barrier will flip from 0 to 1. We create indices of the total number of legal barriers in each country based on three broad categories of laws: barriers to entry, funding, and advocacy. We also sum each of these indices to create an overall count of legal barriers.

- **Barriers to advocacy** (2 possible barriers): (1) NGOs are restricted from engaging in political activities, and (2) advocacy restrictions differ if the NGO receives foreign funding
- **Barriers to entry** (3 possible barriers): (1) registration is burdensome, (2) NGOs are not allowed to appeal if denied registration, and (3) entry requirements differ if the NGO receives foreign funding
- **Barriers to funding** (5 possible barriers): (1) NGOs need prior approval to receive foreign funding, (2) NGOs are required to channel foreign funding to state-owned banks, (3) NGOs face additional restrictions if receiving foreign funding, and (4) NGOs are restricted or (5) prohibited from receiving foreign funding

```{r barriers, fig.width=6, fig.height=6, fig.asp=NULL, fig.cap="Number of legal barriers to NGO activity per country over time (de jure legislation) and average level of CSO repression across democracies and non-democracies (de facto implementation)"}
regulations <- tribble(
    ~question, ~barrier,       ~question_clean,                  ~ignore_in_index, ~question_display,
    "q1a",     "association",  "const_assoc",                    TRUE,             "Constitutional associational rights",
    "q1b",     "association",  "political_parties",              TRUE,             "Citizens form political parties",
    "q4b",     "advocacy",     "ngo_politics_intimidation",      TRUE,             "NGOs intimidated from politics",
    "q4c",     "advocacy",     "ngo_politics_foreign_fund",      FALSE,            "Political barriers differentXXXif foreign funds involved",
    "q2a",     "entry",        "ngo_register",                   TRUE,             "NGO registration required",
    "q2b",     "entry",        "ngo_register_burden",            FALSE,            "NGO registration burdensome",
    "q2c",     "entry",        "ngo_register_appeal",            FALSE,            "NGO registration appealXXXnot allowed",
    "q2d",     "entry",        "ngo_barrier_foreign_funds",      FALSE,            "Registration barriers differentXXXif foreign funds involved",
    "q3a",     "funding",      "ngo_disclose_funds",             TRUE,             "Funds must be disclosed",
    "q3b",     "funding",      "ngo_foreign_fund_approval",      FALSE,            "Prior approval requiredXXXfor foreign funds",
    "q3c",     "funding",      "ngo_foreign_fund_channel",       FALSE,            "Foreign funds channeledXXXthrough government",
    "q3d",     "funding",      "ngo_foreign_fund_restrict",      FALSE,            "Foreign funds restricted",
    "q3e",     "funding",      "ngo_foreign_fund_prohibit",      FALSE,            "Foreign funds prohibited",
    "q3f",     "funding",      "ngo_type_foreign_fund_prohibit", FALSE,            "Foreign funds prohibitedXXXfor some types of NGOs",
    "q4a",     "advocacy",     "ngo_politics",                   FALSE,            "NGOs restricted from politics"
) %>% 
  filter(!ignore_in_index) %>% 
  mutate(barrier_group = paste0("Barriers to ", str_to_lower(barrier))) %>% 
  select(barrier_group, barrier = question_clean, barrier_display = question_display)

df_barriers <- canary %>% 
  group_by(gwcode, year) %>% 
  summarize(across(one_of(regulations$barrier), ~. > 0)) %>% 
  group_by(year) %>% 
  summarize(across(-gwcode, ~sum(., na.rm = TRUE))) %>% 
  ungroup() %>% 
  pivot_longer(names_to = "barrier", values_to = "value", -year) %>% 
  left_join(regulations, by = "barrier") %>% 
  mutate(barrier_display = str_replace(barrier_display, "XXX", "\n")) %>%
  arrange(year, desc(value)) %>% 
  mutate(barrier_display = fct_inorder(barrier_display, ordered = TRUE))

num_countries <- length(unique(canary$gwcode))

entry_plot <- ggplot(filter(df_barriers, 
                            barrier_group == "Barriers to entry"), 
                     aes(x = year, y = value, 
                         color = barrier_display)) +
  geom_line(size = 1.25) +
  expand_limits(y = c(0, 90)) +
  scale_y_continuous(sec.axis = sec_axis(~ . / num_countries,
                                         labels = percent_format(accuracy = 1)),
                     expand = c(0, 0)) +
  scale_color_viridis_d(option = "plasma", end = 0.92, name = NULL) +
  coord_cartesian(xlim = c(1990, 2015), ylim = c(0, 90)) +
  guides(color = guide_legend(nrow = 2)) +
  labs(x = NULL, y = "Number of countries") +
  theme_ngo(9) + 
  theme(legend.justification = "left",
        legend.key.height = unit(0.7, "lines"),
        legend.key.width = unit(1, "lines")) +
  facet_wrap(vars(barrier_group))

funding_plot <- ggplot(filter(df_barriers, 
                              barrier_group == "Barriers to funding"), 
                       aes(x = year, y = value, 
                           color = barrier_display,
                           linetype = barrier_display)) +
  geom_line(size = 1.25) +
  scale_y_continuous(sec.axis = sec_axis(~ . / num_countries,
                                         labels = percent_format(accuracy = 1)),
                     expand = c(0, 0)) +
  scale_color_viridis_d(option = "plasma", end = 0.92, name = NULL) +
  scale_linetype_manual(values = c("solid", "21", "solid", "21", "solid"), name = NULL) +
  coord_cartesian(xlim = c(1990, 2015), ylim = c(0, 45)) +
  guides(color = guide_legend(nrow = 3),
         linetype = guide_legend(nrow = 3)) +
  labs(x = NULL, y = "Number of countries") +
  theme_ngo(9) + 
  theme(legend.justification = "left",
        legend.key.height = unit(0.7, "lines"),
        legend.key.width = unit(1, "lines")) +
  facet_wrap(vars(barrier_group))

advocacy_plot <- ggplot(filter(df_barriers, 
                               barrier_group == "Barriers to advocacy"), 
                        aes(x = year, y = value, 
                            color = barrier_display)) +
  geom_line(size = 1.25) +
  scale_y_continuous(sec.axis = sec_axis(~ . / num_countries,
                                         labels = percent_format(accuracy = 1)),
                     expand = c(0, 0)) +
  scale_color_viridis_d(option = "plasma", end = 0.8, name = NULL) +
  coord_cartesian(xlim = c(1990, 2015), ylim = c(0, 60)) +
  guides(color = guide_legend(nrow = 2)) +
  labs(x = NULL, y = "Number of countries",
       legend.key.size = unit(0.7, "lines")) +
  theme_ngo(9) + 
  theme(legend.justification = "left",
        legend.key.height = unit(0.7, "lines"),
        legend.key.width = unit(1, "lines")) +
  facet_wrap(vars(barrier_group))

df_csreprss_plot <- canary %>%
  group_by(year, autocracy = v2x_regime <= 1) %>%
  nest() %>% 
  mutate(cis = data %>% map(~ mean_cl_normal(.$v2csreprss))) %>% 
  unnest(cis) %>% 
  ungroup() %>% 
  filter(!is.na(autocracy)) %>% 
  mutate(fake_facet_title = "CSO repression",
         autocracy = factor(autocracy, 
                            labels = c("Democracies (Regimes of the World > 1)",
                                       "Non-democracies (Regimes of the World ≤ 1)"), 
                            ordered = TRUE))

csreprss_plot <- ggplot(df_csreprss_plot, aes(x = year, y = y)) +
  geom_ribbon(aes(ymin = ymin, ymax = ymax, fill = autocracy), alpha = 0.2,) +
  geom_line(aes(color = autocracy), size = 1.25) +
  scale_color_viridis_d(option = "plasma", end = 0.8, name = NULL) +
  scale_fill_viridis_d(option = "plasma", end = 0.8, name = NULL) +
  coord_cartesian(xlim = c(1990, 2015)) +
  guides(color = guide_legend(nrow = 2)) +
  labs(y = "Average CSO repression", x = NULL) +
  theme_ngo(9) +
  theme(legend.justification = "left",
        legend.key.height = unit(0.7, "lines"),
        legend.key.width = unit(1, "lines")) +
  facet_wrap(vars(fake_facet_title))

barriers_summary <- 
  ((advocacy_plot | entry_plot) / 
     (funding_plot | csreprss_plot)) &
  theme(legend.text = element_text(size = rel(0.7)),
        axis.title.y = element_text(margin = margin(r = 3)),
        legend.box.margin = margin(t = -0.5, unit = "lines"))

barriers_summary
```

To measure de facto civil society repression in a country, we use the Varieties of Democracy's (V-Dem) CSO repression indicator (`v2csreprss` in @vdem-v10), which captures the extent to which governments harass, deter, liquidate, and arrest members of civil society organizations. This variable captures repression using an ordinal scale, defining civil society repression as severe, substantial, moderate, weak, or absent. This ordinal scale is then aggregated with a Bayesian item response theory measurement model and converted to a continuous scale that ranges from roughly −3 to 3, with higher values indicating less repression.

Figure \@ref(fig:barriers) shows each of these types of legal barriers per country over time, as well as average levels of de facto civil society repression by regime type. While anti-NGO laws have increased steadily since 1990—particularly laws related to foreign funding—there has not necessarily been a corresponding rise in de facto civil society repression, which has remained fairly constant across different regime types. This divergence is likely indicative of the split between de jure laws and their de facto implementation—a rise in laws is not necessarily tied to direct CSO repression.

## Measuring repression

To see how both de jure and de facto civil society restrictions influence general human rights repression, we use four different quantitative measures of repression, each of which capture different dimensions of political and civil violence. First, following @GohdesCarey:2017, who use the killing of journalists to predict worsening human rights conditions, we use the Political Terror Scale (PTS), which assigns countries to one of five possible levels of repression each year [@GibneyCornettWood:2019], with Level 5 representing widespread political violence—including arrests, imprisonment, political murders, disappearances, and torture. As seen in the top panel of Figure \@ref(fig:repression), the number of countries at Level 5 has decreased since the 1990s, while most countries are coded as Levels 2 and 3 (moderate levels of political terror).

The five levels of the PTS are helpful for conceptualizing broader trends in human rights abuses. However, these categories are not always granular enough to model or predict a country's level of political terror. Additionally, the PTS—along with most other measures of human rights abuses—does not account for the systematic changes in the way monitoring agencies have gathered information about human rights violations over time. Accordingly, we use latent human rights scores as an alternative measure of underlying repression [@Fariss:2014; @FarissKenwickReuning:2020]. This variable is generated from a Bayesian measurement model that incorporates information about human rights abuses from a variety of published repression indices, including the PTS. Because the resulting measure is continuous, it picks up on more minute changes in national-level human rights. Moreover, this variable accounts for changing standards of accountability, which has led to the production of more stringent human rights reports as monitors look harder for abuse, look in more places for abuse, and classify more acts as abuse. When accounting for these changes, the bottom panel of Figure \@ref(fig:repression) demonstrates the global improvements in average latent human rights values over time.

```{r repression, fig.width=6.5, fig.height=4.5, fig.asp=NULL, fig.cap="Key measures of repression over time"}
pts_plot <- canary %>% 
  filter(!is.na(PTS_factor)) %>% 
  group_by(PTS_factor, year) %>% 
  summarize(count = n()) %>% 
  mutate(fake_facet_title = "Political Terror Scale") %>% 
  ggplot(aes(x = year, y = count, color = PTS_factor)) +
  geom_line(size = 1) +
  scale_color_viridis_d(option = "rocket", end = 0.85, direction = -1) +
  coord_cartesian(xlim = c(1990, 2015)) +
  guides(color = guide_legend(title = NULL)) +
  labs(y = "Number of countries", x = NULL, caption = "(Higher PTS levels = more repression)") +
  theme_ngo(9) +
  theme(legend.justification = "left",
        legend.key.height = unit(0.7, "lines"),
        legend.key.width = unit(1, "lines"),
        plot.caption = element_text(margin = margin(t = -1, unit = "lines"))) +
  facet_wrap(vars(fake_facet_title))

df_lhr_plot <- canary %>% 
  as_tibble() %>% 
  select(year, latent_hr_mean) %>% 
  group_by(year) %>% 
  nest() %>% 
  mutate(cis = data %>% map(~ mean_cl_normal(.$latent_hr_mean))) %>% 
  unnest(cis) %>% 
  ungroup() %>% 
  mutate(fake_facet_title = "Latent human rights")

lhr_plot <- ggplot(df_lhr_plot, aes(x = year, y = y)) +
  geom_ribbon(aes(ymin = ymin, ymax = ymax), alpha = 0.2, fill = "#FF4136") +
  geom_line(size = 1, color = "#FF4136") +
  coord_cartesian(xlim = c(1990, 2015)) +
  labs(y = "Average value", x = NULL, caption = "(Higher values = less repression)") +
  theme_ngo(9) +
  theme(legend.justification = "left") +
  facet_wrap(vars(fake_facet_title))

(pts_plot / lhr_plot) &
  theme(legend.text = element_text(size = rel(0.7)),
        axis.title.y = element_text(margin = margin(r = 3)),
        legend.box.margin = margin(t = -0.5, unit = "lines"))
```

## Modeling approach

Since we look at whether anti-NGO laws and their implementation can serve as an early warning signal for broader repression, we are primarily interested in *predicting* repression and not isolating the causal effect of civil society repression on general human rights abuses. We do not include a complete set of covariates to remove confounding relationships, but instead use a more parsimonious set of explanatory variables that strongly predict repression. Following @GohdesCarey:2017, we control for a country's level of repression in the previous year (either PTS score or latent human rights value, depending on the model), its level of democracy as measured by V-Dem's polyarchy index, its logged GDP per capita (measured by the UN), the percent of GDP attributed to trade (measured by the UN), and an indicator for whether the country witnessed more than 25 battle-related deaths (measured by @ucdp2020; @GleditschWallensteenEriksson:2002). Because changes in the human rights environment take time to be reflected by these larger indices, we model the effect of lagged explanatory variables on repression one year in the future. We also include the lag of civil society laws or repression to account for serial correlation.

We extend @GohdesCarey:2017's previous modeling approach by accounting for country-level variation in repression with multilevel modeling and we include random effects for each country. Accounting for country differences this way results in intraclass correlation coefficients of around 0.7 for each of our three outcome variables, which means that the country-based structure of the data explains more than 70% of the variation in outcome, resulting in model estimates that are arguably more precisely measured. We run multiple models for each of our empirical expectations, based on different combinations of our outcomes and key explanatory variables:

- **Expectation 1~a~**: Anti-NGO laws (total; advocacy, entry, funding separately) signal worsening political terror
- **Expectation 1~b~**: Anti-NGO laws signal decreasing latent human rights respect
- **Expectation 2~a~**: Civil society repression signals worsening political terror
- **Expectation 2~b~**: Civil society repression signals decreasing latent human rights respect

We generate predictions with two families of Bayesian regression models. Because the PTS is measured as an ordered category, we use ordered logistic regression to predict future values of the scale. We predict the continuous latent human rights values with Gaussian models. The appendix contains complete details of our modeling approach. We use median values from the models' posterior distributions as our point estimates and provide credible intervals using the 95% highest posterior density. We declare an effect statistically "significant" if the posterior probability of being different from zero (i.e. above or below) exceeds 0.95.

# Results and analysis

We present the results of our models in two stages. First, to see how minor shifts in civil society laws and repression influence future human rights, we examine the marginal effects of our main explanatory variables when all other model covariates are held constant. These effects are by no means causal—rather, they demonstrate how much variation in predicted human rights is possible as the legal environment for civil society improves or deteriorates within a typical country. Second, we present predictive diagnostics for each of our models to examine the improvements in predictions. The results of the ordered logistic regression models can be unwieldy to interpret with plain numbers due to varying intercepts and thresholds between PTS levels. Accordingly, we present graphical results for all our models wherever possible and include complete tables of model results in the appendix (see Tables \@ref(tab:tbl-models-pts) and \@ref(tab:tbl-models-lhr)). For ease in comparing the count of NGO laws with civil society repression, we reverse the x-axis in the marginal effects plots for our second empirical expectation so that increasing the civil society repression represents *worse* repression rather than less.

## Marginal results

### How does repression change following increases in NGO laws?

In general, when holding all other covariates constant, the passage of an additional anti-NGO legal barrier is generally associated with a higher probability of more severe level of political terror in the following year. The probability of seeing different levels of political terror varies across different possible counts of NGO laws (see Figure \@ref(fig:plot-mfx-e1a)). In the absence of formal legal barriers, Level 2 of the PTS—an environment where political imprisonment, torture, and beatings are rare—is the most likely category of political terror across all models, appearing a predicted 60% of the time. Level 1, where a country is under secure rule of law, is the second most likely outcome at slightly under 40%. The other categories of the PTS are extremely rare when there are no anti-NGO laws—as such, we collapse them into a single category.

```{r plot-mfx-e1a, fig.width=6.5, fig.height=5, fig.asp=NULL, fig.cap="Marginal effects of increasing anti-NGO legal barriers on the probability of specific levels of political terror"}
mfx_e1a_pts %>% 
  mutate(x_rounded = floor(effect1__)) %>% 
  # Collapse 3-5 to just one level, since they're all tiny
  mutate(effect2__ = fct_collapse(effect2__, "3+" = c("3", "4", "5"))) %>% 
  group_by(x_rounded, effect2__, plot_var_nice) %>% 
  slice(1) %>% 
  # Add up these values for the collapsed 3-5 results
  summarize(across(c(estimate__, lower__, upper__), sum)) %>% 
  ungroup() %>% 
  ggplot(aes(x = x_rounded, y = estimate__, color = effect2__)) +
  geom_line(aes(group = effect2__), size = 0.5, color = "grey90") +
  geom_pointrange(aes(ymin = lower__, ymax = upper__),
                  # position = position_dodge(width = 0.2),
                  fatten = 1, size = 1) +
  scale_x_continuous(breaks = 0:9) +
  scale_y_continuous(labels = percent_format(accuracy = 1)) +
  scale_color_viridis_d(option = "rocket", begin = 0.1, end = 0.85, direction = -1) +
  guides(color = guide_legend(override.aes = list(size = 0.5, linetype = 0))) +
  labs(x = "Count of legal barriers\n",
       y = "Predicted probability of category",
       color = "Predicted Political Terror Scale", fill = "Predicted Political Terror Scale",
       title = "Political Terror Scale") + 
  facet_wrap(vars(plot_var_nice), scales = "free_x") +
  theme_ngo() +
  theme(panel.grid.major.x = element_blank())
```

As new laws are added, though, the distribution of probabilities for each of the PTS levels shifts. Adding new barriers to advocacy, entry, and funding decreases the likelihood of seeing Level 1 and increases the probability of a country turning to either Level 2 or Level 3 (with extensive political imprisonment and trial-free detention). The marginal effects of all forms of NGO laws results in this turn from Level 1 to Levels 2 or 3+, but this switching effect is strongest following additional barriers to advocacy. Anti-NGO laws do not predict more dramatic shifts in the PTS. Across all counts of legal barriers, the overall probability of seeing Levels 4 and 5 is minuscule, and the passage of additional laws does little to shift that probability. This is not unexpected however—the PTS has a natural ceiling at Level 5. Specifically, in addition to measuring the scope and intensity of government violence, the PTS scale also looks at the proportion of population targeted for abuse. Countries that receive a 4 and 5 are rare, as 4 denotes that murders, disappearances, and torture are a common part of life, while 5 denotes that physical abuses are routinely perpetrated against citizens not involved in politics, and includes cases where governments engage in "scorched earth" counterinsurgency policies or summary executions based on class or ethnic affiliation [@GibneyCornettWood:2019]. The majority of movement across PTS categories should thus occur between Levels 1–3.

Figure \@ref(fig:plot-mfx-e1b) shows the marginal effects of additional NGO legislation on latent human rights scores. Across all four models for different categories of legal barriers, holding all other values constant, the predicted level of human rights respect decreases. Adding a new NGO law is associated with a -0.013-point decline in the level of respect for human rights in the following year, on average. The marginal effect is larger (-0.02) for additional barriers to advocacy and entry, which represent more burdensome restrictions than limits to funding.

At a marginal level, therefore, adding additional anti-NGO laws in countries with low levels of political terror could signal that higher levels of political repression are more likely in the future. Limited repression remains the most likely possible outcome, but the chance of seeing robust rule of law decreases rapidly as more laws are passed, all else equal. This pattern holds when looking at a more continuous measure of human rights protection—countries with a low count of anti-NGO legal barriers tend to have higher predicted values of latent human rights scores, and average protection against abuses decreases as more laws are added.

```{r plot-mfx-e1b, fig.width=7.5, fig.height=4, fig.asp=NULL, fig.cap="Marginal effects of increasing anti-NGO legal barriers on latent human rights scoress"}
# TODO: Andrew: Add more specific coefficients and percent changes to the prose above
out_mfx_e1b <- mfx_e1d_lhr %>% 
  mutate(x_rounded = floor(effect1__)) %>% 
  group_by(x_rounded, plot_var_nice) %>% 
  slice(1) %>% 
  ungroup() %>% 
  mutate(plot_var_nice = fct_relabel(
    plot_var_nice, ~str_to_sentence(str_remove(., "Barriers to ")))) %>% 
  ggplot(aes(x = x_rounded, y = estimate__, color = plot_var_nice)) +
  geom_line(size = 0.5, color = "grey90") +
  geom_pointrange(aes(ymin = lower__, ymax = upper__),
                  fatten = 1, size = 1) +
  scale_x_continuous(breaks = 0:9) +
  scale_color_viridis_d(option = "plasma", end = 0.8) +
  guides(color = "none") +
  labs(x = "Count of legal barriers",
       y = "Predicted latent human rights value",
       title = "Latent human rights",
       caption = "(Higher values = better protection)") +
  facet_grid(cols = vars(plot_var_nice), scales = "free_x", space = "free_x") +
  theme_ngo() +
  theme(panel.grid.major.x = element_blank())

out_mfx_e1b
```

### How does general repression change following civil society repression?

The weaker predictive power of formal NGO laws is likely attributable to fact that de jure NGO legislation may take time to implement—many states need to first establish institutions that oversee the operation and funding of these groups or incorporate threatening NGOs into the regime apparatus [@Naim:2007]. Further, states that engage in political repression can do so with or without legal justification. De facto civil society repression, on the other hand, represents the actual implementation of anti-NGO strategies—both legal and extralegal—and might be more indicative of future human rights abuses. The results from our second set of models, using V-Dem;s civil society repression index as the key predictor, confirm this.

The effect of de facto civil society repression on predicted values of the PTS mirrors what we found previously with our first empirical expectation. As seen in Figure \@ref(fig:plot-mfx-e2a), at the lowest levels of repression, Levels 1 and 2 are the most common predicted outcomes and are each equiprobable at roughly 50%. As de facto CSO repression increases, though, the probability of seeing Level 1 decreases and is replaced with either Level 2 or Level 3. More severe levels of political terror remain exceptionally improbable at any level of preceding civil society repression—we again collapse these into a single category.

```{r plot-mfx-e2a, fig.width=6.5, fig.height=5, fig.asp=NULL, fig.cap="Marginal effects of changing levels of civil society repression on the probability of specific levels of political terror", out.width="75%"}
out_mfx_e2a <- mfx_e2a_pts %>% 
  mutate(effect2__ = fct_collapse(effect2__, "3+" = c("3", "4", "5"))) %>% 
  group_by(effect1__, effect2__, plot_var_nice) %>% 
  summarize(across(c(estimate__, lower__, upper__), sum)) %>% 
  ungroup() %>% 
  ggplot(aes(x = effect1__, y = estimate__, color = effect2__, fill = effect2__)) +
  geom_ribbon(aes(ymin = lower__, ymax = upper__), alpha = 0.4, color = NA) +
  geom_line(size = 1) +
  scale_x_reverse() +
  scale_y_continuous(labels = percent_format(accuracy = 1)) +
  scale_color_viridis_d(option = "rocket", begin = 0.1, end = 0.75, direction = -1) +
  scale_fill_viridis_d(option = "rocket", begin = 0.1, end = 0.9, direction = -1) +
  labs(x = "Civil society repression",
       y = "Predicted probability of category",
       color = "Political Terror Scale", fill = "Political Terror Scale",
       title = "Political Terror Scale",
       caption = "(x-axis reversed; greater repression on the right)") +
  theme_ngo() +
  theme(legend.box.margin = margin(t = 0.5, unit = "lines"))
out_mfx_e2a
```

```{r lhr-trends, include=FALSE, warning=FALSE}
lhr_csreprss_stats <- m_lhr_v2csreprss %>% 
  tidy(effects = "fixed") %>% 
  mutate(term = janitor::make_clean_names(term)) %>%
  split(.$term)

lhr_csreprss_diffs <- m_lhr_v2csreprss %>% 
  emmeans(~ v2csreprss + gwcode, var = "v2csreprss", 
           at = list(gwcode = 999,
                     v2csreprss = c(0, -1)),
           epred = TRUE, re_formula = NULL,
           allow_new_levels = TRUE, sample_new_levels = "gaussian") %>% 
  tidy() %>% 
  mutate(pct_change = (estimate - lag(estimate)) / lag(estimate),
         id = janitor::make_clean_names(v2csreprss)) %>% 
  split(.$id)
```

De facto civil society repression has a noticeable effect on latent human rights values (see Figure \@ref(fig:plot-mfx-e2b)). When holding all other model parameters at their typical values, increased civil society repression predicts a strong decrease in respect for broader human rights: a one unit change in civil society repression is associated with a `r round(lhr_csreprss_stats$v2csreprss$estimate, 2)`-unit change in overall latent human rights. To illustrate the magnitude of this effect, a typical simulated country with a civil society repression index of 0 is predicted to have a latent human rights score of `r round(lhr_csreprss_diffs$x0$estimate, 2)`. This predicted outcome drops to `r round(lhr_csreprss_diffs$x1$estimate, 2)` with a 1-unit decrease in civil society repression, representing a `r percent_format(accuracy = 1)(abs(lhr_csreprss_diffs$x1$pct_change))` decline in respect for human rights in the following year.

When holding all other factors equal, the de facto implementation of anti-civil society measures thus appear to be a stronger marginal indicator of future repression than the passage of formal de jure laws. In countries with low levels of political terror, it can potentially serve as a weak signal of a more general increase in political violence in the future, but predict more substantial changes in general respect for human rights. The difference in the effects of laws vs. their de facto implementation could also potentially be explained by the time frame of these strategies. De jure laws are a long-term strategy to prevent threats to the regime from coalescing in the first place. For instance, Egypt's Law 84 of 2002—which outlined many anti-NGO restrictions—has been on the books for nearly two decades, but only over the past five years have post-revolutionary regimes used it more broadly [@Ruffner:2015]. As such, the long-term effects of these laws in terms of their predictive power may require a much longer time period under study. In contrast, de facto crackdown on NGOs may signal the need to react sooner due to the immediacy of the threat.

```{r plot-mfx-e2b, fig.width=6.5, fig.height=4, fig.asp=NULL, fig.cap="Marginal effects of changing levels of civil society repression on latent human rights"}
# TODO: Combine with previous plot so there's just one, with fake facet titles
out_mfx_e2d <- mfx_e2d_lhr %>% 
  ggplot(aes(x = effect1__, y = estimate__)) +
  geom_ribbon(aes(ymin = lower__, ymax = upper__), alpha = 0.4, 
              fill = "#FF4136", color = NA) +
  geom_line(size = 1, color = "#FF4136") +
  scale_x_reverse() +
  labs(x = "Civil society repression",
       y = "Predicted latent human rights value",
       title = "Latent human rights", 
       caption = "(x-axis reversed; greater repression on right\nhigher values in y-axis show less violence and better protection of rights)") +
  theme_ngo(8)
out_mfx_e2d
```

## Predictions

```{r eval=FALSE, include=FALSE}
# TODO: Andrew: Streamline this section
```

These predicted probabilities and outcomes are conditional on all other model parameters being held at their average values and show the effect of hypothetically changing a single predictor to different values. They do not show how well these models work with actual data. Following @WardBeger:2017 and @GohdesCarey:2017, we test the predictive power of both sets of our models using out-of-sample prediction. We divide our complete data into a training set and a test set—the training set includes data from all countries from 1990–2010, and the test set includes the last three years of data (2011–2013). We re-run each of our models using our training data and then use the estimated parameters to predict political terror, physical violence, and private civil liberties for the test data. Additionally, we estimate a model for each outcome without predictor variables for civil society laws or repression as a baseline measure of model performance.

```{r correct-not-correct}
# These numbers come from the analysis notebook at 
# https://stats.andrewheiss.com/canary-ngos/04_predictions.html#Prediction_diagnostics
pts_pred_table <- tribble(
  ~Prediction, ~Baseline, ~total, ~advocacy, ~entry, ~funding, ~repression,
  "Wrong",     108,       108,    110,       108,    109,      111,
  "Correct",   377,       377,    375,       377,    376,      374
) %>% 
  rename("Total legal barriers included" = total,
         "Barriers to advocacy included" = advocacy,
         "Barriers to entry included" = entry,
         "Barriers to funding included" = funding,
         "Civil society repression included" = repression)

pts_pred_table_caption <- "Counts of correctly and incorrectly predicted cases across different PTS models"

if (!is_docx) {
  pts_pred_table %>% 
    kbl(align = "lcccccc",
        booktabs = TRUE, linesep = "\\addlinespace",
        caption = pts_pred_table_caption) %>% 
    kable_styling(font_size = ifelse(is_latex, 8.5, NA)) %>%
    kable_styling() %>% 
    column_spec(1, width = "0.07\\\\textwidth") %>% 
    column_spec(2, width = "0.05\\\\textwidth") %>% 
    column_spec(3:7, width = "0.13\\\\textwidth")
} else {
  pts_pred_table %>% 
    hux() %>% 
    set_font_size(10) %>%
    set_all_padding(3) %>% 
    set_bold(1, everywhere) %>%
    set_bottom_border(1, everywhere) %>%
    # set_col_width(1, 0.05) %>%
    # set_col_width(2:7, 0.14) %>% 
    set_width(1) %>% 
    set_wrap(TRUE) %>% 
    huxtable::set_caption(pts_pred_table_caption) %>% 
    huxtable::as_flextable()
}
```

We assess predictive power in a few different ways, depending on the model used to predict the outcome. For the ordinal Political Terror Scale, Table \@ref(tab:correct-not-correct) shows a count of correct and incorrect predictions from each model. Because the physical violence and private civil liberties indexes are predicted with Gaussian models, we measure predictive accuracy with the root mean squared error, provided in Table \@ref(tab:tbl-rmse). In the appendix we include lists of country-year cases that see improved predictions when taking civil society laws and repression into account.

Including civil society restrictions—both de jure and de facto—does very little to improve predictions of political terror. As seen in Table \@ref(tab:correct-not-correct), no model provides more correct predictions than the baseline model, and some models perform worse. Accounting for anti-NGO activities does improve predictions in eight cases (see Table \@ref(tab:tbl-improved-pts)), but it also results in an equal number of incorrect predictions.

```{r tbl-rmse}
# These numbers also come from the analysis notebook at 
# https://stats.andrewheiss.com/canary-ngos/04_predictions.html
tbl_rmse <- tribble(
  ~`Latent human rights values`,  ~RMSE, ~`% change from baseline`,
                  "Baseline", 0.1861944,                       "—",
      "Total legal barriers", 0.1872427,                   "0.56%",
      "Barriers to advocacy", 0.1871498,                   "0.51%",
         "Barriers to entry", 0.1862244,                   "0.02%",
       "Barriers to funding", 0.1879829,                   "0.96%",
  "Civil society repression", 0.1856543,                  "-0.29%"
) %>% 
  mutate(RMSE = round(RMSE, 4))

tbl_rmse_caption <- "Root mean squared error across models predicting latent human rights"

if (!is_docx) {
  tbl_rmse %>% 
    kbl(align = "lcc",
        booktabs = TRUE, linesep = "\\addlinespace",
        caption = tbl_rmse_caption) %>% 
    kable_styling()
} else {
  tbl_rmse %>% 
    hux() %>% 
    set_font_size(10) %>%
    set_all_padding(3) %>% 
    set_align(everywhere, 1, "left") %>% 
    set_align(everywhere, 2:3, "center") %>% 
    set_bold(1, everywhere) %>%
    set_width(1) %>% 
    set_wrap(TRUE) %>% 
    huxtable::set_caption(tbl_rmse_caption) %>% 
    huxtable::as_flextable()
}
```

When looking at the level of physical violence in a country, including de jure and de facto repression does very little to improve predictive accuracy. As Table \@ref(tab:tbl-rmse) shows, accounting for legal barriers and civil society repression only reduces the RMSE by at most 0.4% compared to the baseline model. This is likely related to the lack of improvements in predicting political terror—both indices measure similar phenomena. Similarly, accounting for de facto civil society repression when modeling latent human rights only results in slight changes to the RMSE. However, including civil society-related predictors when modeling private civil liberties does result in small gains in accuracy, reducing the RMSE by as much as 2% compared to the baseline model.

While these gains in predictive power appear negligible, they are somewhat comparable with the findings of previous work. For instance, in their work on the effect of the murder of journalists on predicting future political terror, @GohdesCarey:2017 find that accounting for violence against the media adds five correctly predicted cases of PTS scores and improves predictions in seven cases.

# Conclusion and implications

Strengthening civil society is an important precondition for democratic transition and consolidation. When governments restrict civil society, it reduces these organizations' ability to maintain accountability and protect against abuse. Given these potential knock-on effects, in this paper, we posited that the ongoing global crackdown on NGOs could predict worsening human rights abuses and levels of political and civil rights repression. Using original data on de jure and de facto NGO repression from 1990-2013, we find little support for the predictive power of de jure restriction on political terror, physical violence, and violations of private civil liberties. Marginal predictions show that, all else held equal, adding new anti-NGO laws increases the probability of seeing moderate levels of political terror, but does not change predicted values of physical integrity or civil liberties rights violations. When comparing out-of-sample predictions to a baseline model, including de jure restrictions as predictors does not improve our models' predictive accuracy.

We do, however, find some predictive power when accounting for de facto levels of NGO repression, which includes both legal and extra-legal means of crackdown. As with formal laws, rising NGO repression tends to increase the probability of more moderate political terror. Additionally, this repression is associated with worsening respect for physical integrity rights and civil liberties in the future. In out-of-sample predictions, including civil society restrictions improves model accuracy by 1–2% over the baseline.

Importantly, the lack of predictive power when modeling political violence and the modest improvements in predictive power when modeling civil liberties likely reflects the purpose of NGO restrictions. Encroachments on private civil liberties such as restricting freedom of association are not nearly as violent as events like government-sponsored murders. In authoritarian states in particular, administrative crackdown on NGOs can also be seen as a careful strategy for autocratic survival. Authoritarian leaders must carefully balance domestic and international institutional pressures—including pressure from civil society and NGOs—to remain in power [@Heiss:2019a; @LevitskyWay:2010]. Anti-NGO laws could possibly be a side effect of authoritarian stability-seeking behavior, similar to election tampering, media censorship, cronyism, and rent-seeking, rather than an overt signal of impending serious human rights abuses.

The global shrinking of civil society is thus not likely an early warning sign for future violent repression—states typically signal their future repressive intentions with more violent methods, such as the persecution and murder of activists and the media. Instead, civil society is a possible "canary in the coal mine" for impending violations of other civil rights, such as the freedom of religion, labor rights, property rights, and the freedom of movement.

\clearpage

# Appendix

## Modeling approach

We use Stan [@stan] through R [@r-project] and brms [@brms-jss] to estimate our models. We generate 4 MCMC chains for each model with 2,000 iterations in each chain, 1,000 of which are used for warmup. All chains converge; we assess convergence with visual inspection.

Complete results from all the models, along with posterior predictive checks, goodness-of-fit measures, and prediction diagnostics are all available at a companion website at `anonymized`. Our general modeling approach can be summarized as follows:

\footnotesize

$$
\begin{aligned}
& \textbf{Likelihoods} \\
y_{i, t+1} \sim&\ \operatorname{Ordered Logistic}(y^*_{i, t + 1}, \sigma^2_m) & \text{[PTS]} \\
y_{i, t+1} \sim&\ \mathcal{N}(y^*_{i, t + 1}, \sigma^2_m) & \text{[latent human rights]} \\
\ \\
& \textbf{Parameters} \\
y^*_{i, t + 1} =&\ \alpha_i + \beta_1 \text{Civil society repression}_{i, t} + & \text{[model with covariates]}\\
&\ \beta_2 \text{Civil society repression}_{i, t - 1} + \\
&\ \beta_3 \text{Outcome}_{i, t} + \beta_4 \text{Polyarchy}_{i, t} + \\
&\ \beta_5 \log (\text{GDP per capita}) + \\
&\ \beta_6 \text{Trade as } \% \text{ of GDP}_{i, t} + \\
&\ \beta_7 \text{Armed conflict}_{i, t} + \sigma^2_e \\
\alpha_i =&\ \mathcal{N}(0, \sigma^2_\alpha) & \text{[country-specific intercepts]} \\
\ \\
& \textbf{Priors} \\
\alpha \sim&\ \mathcal{N}(0, 10) & \text{[population intercept]} \\
\beta_{1-7} \sim&\ \mathcal{N}(0, 3) & \text{[population effects]} \\
\sigma^2_m, \sigma^2_e, \sigma^2_\alpha \sim&\ \operatorname{Cauchy}(0, 1) & \text{[sd for model, population,} \\
&&\text{and country error]}
\end{aligned}
$$
\normalsize

To check for robustness, we also ran models using both country and regional random effects, but doing so made no noticeable difference in the results—for the sake of computational efficiency, we thus do not include regional effects.

\newpage

## Improved cases

```{r tbl-improved-pts}
tbl_improved_pts <- tribble(
               ~Country, ~Year,   ~Actual, ~`Baseline prediction`, ~`Total NGO barriers included`, ~`Civil society repression included`,
            "Guatemala", 2013L, "Level 3", "Level 2",                    "Level 3",                          "Level 3",
               "Kosovo", 2012L, "Level 1", "Level 2",                    "Level 1",                          "Level 1",
         "Sierra Leone", 2011L, "Level 3", "Level 2",                    "Level 3",                                "—",
  "Congo - Brazzaville", 2011L, "Level 3", "Level 2",                    "Level 3",                          "Level 3",
              "Somalia", 2013L, "Level 5", "Level 4",                    "Level 5",                                "—",
              "Bahrain", 2011L, "Level 3", "Level 2",                    "Level 3",                          "Level 3",
              "Bahrain", 2013L, "Level 3", "Level 2",                    "Level 3",                          "Level 3",
                "Libya", 2011L, "Level 4", "Level 5",                          "—",                          "Level 4"
  )

tbl_improved_pts_caption <- "Country-year cases where including civil society restrictions improved PTS predictions"

if (!is_docx) {
  tbl_improved_pts %>% 
    kbl(align = "llcccc",
        booktabs = TRUE, linesep = "\\addlinespace",
        caption = tbl_improved_pts_caption) %>% 
    kable_styling(latex_options = c("hold_position")) %>% 
    row_spec(0, bold = TRUE) %>% 
    column_spec(1, width = "0.1\\\\textwidth") %>% 
    column_spec(2, width = "0.05\\\\textwidth") %>% 
    column_spec(3:6, width = "0.16\\\\textwidth")
} else {
  tbl_improved_pts %>% 
    hux() %>% 
    set_font_size(10) %>%
    set_all_padding(3) %>% 
    set_align(everywhere, 1:2, "left") %>% 
    set_align(everywhere, 3:6, "center") %>% 
    set_bold(1, everywhere) %>%
    set_bottom_border(1, everywhere) %>%
    set_width(1) %>% 
    set_wrap(TRUE) %>% 
    huxtable::set_caption(tbl_improved_pts_caption) %>% 
    huxtable::as_flextable()
}
```

\newpage

\blandscape

## Complete model results

\memSingleSmall

```{r tbl-models-pts}
model_names_tbl_pts <- c("Total barriers",
                         "Barriers to advocacy",
                         "Barriers to entry",
                         "Barriers to funding",
                         "Civil society repression")

names(models_paper_pts) <- model_names_tbl_pts

tbl_reg_pts_caption <- "Full results from ordered logistic regression models predicting political terror"

tbl_reg_pts <- modelsummary(models_paper_pts,
                            output = table_format,
                            statistic = "[{conf.low}, {conf.high}]",
                            coef_map = coef_list,
                            gof_omit = "ELPD",
                            escape = FALSE,
                            caption = tbl_reg_pts_caption,
                            notes = c(model_note),
                            longtable = TRUE)

if (!is_docx) {
  tbl_reg_pts %>% 
    kable_styling() %>% 
    add_header_above(header = c(" " = 1, "Outcome in t + 1" = 5))
    # kable_styling(font_size = ifelse(is_latex, 7, NA)) %>%
    # column_spec(1, width = "0.75in", latex_valign = "b") %>%
    # column_spec(2:6, width = "1.3in", latex_valign = "b")
    # column_spec(1, width = "0.10\\\\textwidth") %>%
    # column_spec(2:6, width = "0.14\\\\textwidth")
} else {
  tbl_reg_pts %>% 
    set_font_size(10) %>% 
    set_all_padding(1) %>% 
    set_align(everywhere, 2:6, "center") %>% 
    set_bold(1, everywhere) %>% 
    # set_col_width(1, 0.1) %>% 
    # set_col_width(2:6, 0.14) %>% 
    set_width(1) %>% 
    set_wrap(TRUE) %>% 
    huxtable::set_caption(tbl_reg_pts_caption) %>% 
    huxtable::as_flextable()
}
```

\clearpage

```{r tbl-models-lhr}
model_names_tbl_lhr <- c("Total barriers",
                         "Barriers to advocacy",
                         "Barriers to entry",
                         "Barriers to funding",
                         "Civil society repression")

names(models_paper_lhr) <- model_names_tbl_lhr

tbl_reg_lhr_caption <- "Full results from Gaussian regression models predicting latent human rights"

tbl_reg_lhr <- modelsummary(models_paper_lhr,
                            output = table_format,
                            statistic = "[{conf.low}, {conf.high}]",
                            coef_map = coef_list,
                            gof_omit = "ELPD",
                            escape = FALSE,
                            caption = tbl_reg_lhr_caption,
                            notes = c(model_note),
                            longtable = TRUE)

if (!is_docx) {
  tbl_reg_lhr %>% 
    kable_styling() %>% 
    add_header_above(header = c(" " = 1, "Outcome in t + 1" = 5))
    # kable_styling(font_size = ifelse(is_latex, 7.5, NA)) %>% 
    # column_spec(1, width = "0.10\\\\textwidth") %>% 
    # column_spec(2:6, width = "0.14\\\\textwidth")
} else {
  tbl_reg_lhr %>% 
    set_font_size(10) %>% 
    set_all_padding(1) %>% 
    set_align(everywhere, 2:6, "center") %>% 
    set_bold(1, everywhere) %>% 
    set_width(1) %>% 
    set_wrap(TRUE) %>% 
    huxtable::set_caption(tbl_reg_lhr_caption) %>% 
    huxtable::as_flextable()
}
```

\clearpage

\elandscape

\memSingle