### 2021-11-15 manuscript/appendix.Rmd
Move appendix to separate document

---
title: "NGO Repression as a Predictor of Worsening Human Rights Abuses"
short-title: "NGO Repression and Human Rights"
author:
- name: Suparna Chaudhry
  affiliation: Lewis & Clark College
  email: schaudhry@lclark.edu
  url: http://www.suparnachaudhry.com/
- name: Andrew Heiss
  affiliation: Georgia State University
  email: aheiss@gsu.edu
  url: https://www.andrewheiss.com/
date: "July 20, 2021"
published: Under review
code-repo: "Access the code and full analysis notebook at <https://stats.andrewheiss.com/canary-ngos/>"
abstract: >-
  An increasing number of countries have recently cracked down on non-governmental organizations (NGOs). Much of this crackdown is sanctioned by law and represents a bureaucratic form of repression that could indicate more severe human rights abuses in the future. This is especially the case for democracies, which unlike autocracies, may not aggressively attack civic space. We explore if crackdowns on NGOs predict broader human rights repression. Anti-NGO laws are among the most subtle means of repression and attract lesser domestic and international condemnation compared to the use of violence. Using original data on NGO repression, we test whether NGO crackdown is a predictor of political terror, and violations of physical integrity rights and civil liberties. We find that while formal de jure laws provide little information in predicting future repression, their patterns of implementation, or de facto civil society repression, predicts worsening respect for physical integrity rights and civil liberties.
# keywords: civil society, NGOs, repression, human rights, political terror, civil liberties
word-count: "7,159"
thanks: >-
  We would like to thank participants of the 2021 International Studies Association (ISA) convention, and the "Human Rights on the Edge: The Future of International Human Rights Law and Practice" workshop hosted by the Global Human Rights Hub at Arizona State University for their comments and feedback.
reference-section-title: References
bibliography: bibliography.bib
toc: false
title-page: true
endnotes: true
endfloat: false
mainfont: Cochineal
sansfont: Inter
author-note: >-
  Suparna Chaudhry, PhD, Assistant Professor of International Affairs, Department of International Affairs, Lewis \& Clark College; Andrew Heiss, PhD, Assistant Professor of Public Management, Andrew Young School of Policy Studies, Georgia State University. 
correspondence: >-
  Correspondence concerning this article should be addressed to Andrew Heiss, PhD, Department of Public Management and Policy, Andrew Young School of Policy Studies, Georgia State University, 14 Marietta Street NW, Suite 357, Atlanta, GA 30303. Contact: aheiss@gsu.edu
editor_options: 
  chunk_output_type: console
---

# Appendix

## Modeling approach

We use Stan [@stan] through R [@r-project] and brms [@brms-jss] to estimate our models. We generate 4 MCMC chains for each model with 2,000 iterations in each chain, 1,000 of which are used for warmup. All chains converge; we assess convergence with visual inspection.

To check for robustness, we also ran models using both country and regional random effects, but doing so made no noticeable difference in the results—for the sake of computational efficiency, we thus do not include regional effects.

Complete results from all the models, along with posterior predictive checks, goodness-of-fit measures, and prediction diagnostics are all available at a companion statistical analysis notebook at <https://doi.org/10.17605/OSF.IO/MTR6X>. Our general modeling approach can be summarized as follows:

### Equation for PTS models

$$
\begin{aligned}
& \textbf{Likelihood for ordinal outcome} \\
& \text{Political Terror Scores (PTS)} \\
y_{i, t+1} \sim&\ \operatorname{Categorical}(p_{i, t + 1}) \\
\ \\
& \textbf{Parameters} \\
\operatorname{logit}(p_i) =&\ \alpha_{j[i]} - \phi_{k, t + 1} \\
\phi_{k, t + 1} =&\ \beta_1 \text{PTS}_{i, t} + \beta_2 \text{PTS}_{i, t - 1} + \\
&\ \beta_3 \text{Outcome}_{i, t} + \beta_4 \text{Polyarchy}_{i, t} + \\
&\ \beta_5 \log (\text{GDP per capita}) + \\
&\ \beta_6 \text{Trade as } \% \text{ of GDP}_{i, t} + \\
&\ \beta_7 \text{Armed conflict}_{i, t} \\ \\
\alpha_j \sim&\ \mathcal{N} (\mu_{\alpha_j}, \sigma_{\alpha_j}), \text{ for country } j \text{ in } 1 .. J & \text{[country-specific intercepts]} \\ \\
\ \\
& \textbf{Priors} \\
\mu_{\alpha_j} \sim&\ \mathcal{N}(0, 10) & \text{[country-specific intercepts]} \\
\beta_{1-7} \sim&\ \mathcal{N}(0, 3) & \text{[population effects]} \\
\sigma_{\alpha_j} \sim&\ \operatorname{Cauchy}(0, 1) & \text{[sd for population \& country]}
\end{aligned}
$$

### R code for PTS models

The actual R code for this model is included in the replication code at <https://doi.org/10.17605/OSF.IO/MTR6X>. This is a simplified representation of the **brms** [@brms-jss] model code.

```{r echo=TRUE, eval=FALSE}
# Ordinal logistic regression for political terror
# (replace `cs_repression` with different measures of de jure and 
# de facto civil society repression: barriers_total, advocacy, entry, 
# funding, and v2csreprss)
brm(
  bf(PTS_lead1 ~ cs_repression + cs_repression_lag1 +
       PTS + v2x_polyarchy + gdpcap_log +
       un_trade_pct_gdp + armed_conflict + (1 | gwcode)),
  family = cumulative(),
  prior = c(set_prior("normal(0, 10)", class = "Intercept"),
            set_prior("normal(0, 3)", class = "b"),
            set_prior("cauchy(0, 1)", class = "sd")),
  ...
)
```

### Equation for latent human rights models

$$
\begin{aligned}
& \textbf{Likelihood for continuous outcome} \\
& \text{Latent human rights (LHR) values} \\
y_{i, t+1} \sim&\ \mathcal{N}(y^*_{i, t + 1}, \sigma_i) \\
\ \\
& \textbf{Parameters} \\
y^*_{i, t + 1} =&\ \alpha_{j[i]} + \beta_1 \text{LHR}_{i, t} +  \beta_2 \text{LHR}_{i, t - 1}\\
&\ \beta_3 \text{Outcome}_{i, t} + \beta_4 \text{Polyarchy}_{i, t} + \\
&\ \beta_5 \log (\text{GDP per capita}) + \\
&\ \beta_6 \text{Trade as } \% \text{ of GDP}_{i, t} + \\
&\ \beta_7 \text{Armed conflict}_{i, t}\\ \\
\alpha_j \sim&\ \mathcal{N} (\mu_{\alpha_j}, \sigma_{\alpha_j}), \text{ for country } j \text{ in } 1 .. J & \text{[country-specific intercepts]} \\ \\
& \textbf{Priors} \\
\mu_{\alpha_j} \sim&\ \mathcal{N}(0, 10) & \text{[country-specific intercepts]} \\
\beta_{1-7} \sim&\ \mathcal{N}(0, 3) & \text{[population effects]} \\
\sigma_i, \sigma_{\alpha_j} \sim&\ \operatorname{Cauchy}(0, 1) & \text{[sd for population \& country]}
\end{aligned}
$$

### R code for latent human rights models

The actual R code for this model is included in the replication code at <https://doi.org/10.17605/OSF.IO/MTR6X>. This is a simplified representation of the **brms** [@brms-jss] model code.

```{r echo=TRUE, eval=FALSE}
# Gaussian regression for latent human rights
# (replace `cs_repression` with different measures of de jure and 
# de facto civil society repression: barriers_total, advocacy, entry, 
# funding, and v2csreprss)
brm(
  bf(latent_hr_mean_lead1 ~ cs_repression + cs_repression_lag1 +
       latent_hr_mean + v2x_polyarchy + gdpcap_log +
       un_trade_pct_gdp + armed_conflict + (1 | gwcode)),
  family = gaussian(),
  prior = c(set_prior("normal(0, 10)", class = "Intercept"),
            set_prior("normal(0, 3)", class = "b"),
            set_prior("cauchy(0, 1)", class = "sd")),
  ...
)
```

\newpage

## Improved cases

```{r tbl-improved-pts}
tbl_improved_pts <- tribble(
               ~Country, ~Year,   ~Actual, ~`Baseline prediction`, ~`Total NGO barriers included`, ~`Civil society repression included`,
            "Guatemala", 2013L, "Level 3", "Level 2",                    "Level 3",                          "Level 3",
               "Kosovo", 2012L, "Level 1", "Level 2",                    "Level 1",                          "Level 1",
         "Sierra Leone", 2011L, "Level 3", "Level 2",                    "Level 3",                                "—",
  "Congo - Brazzaville", 2011L, "Level 3", "Level 2",                    "Level 3",                          "Level 3",
              "Somalia", 2013L, "Level 5", "Level 4",                    "Level 5",                                "—",
              "Bahrain", 2011L, "Level 3", "Level 2",                    "Level 3",                          "Level 3",
              "Bahrain", 2013L, "Level 3", "Level 2",                    "Level 3",                          "Level 3",
                "Libya", 2011L, "Level 4", "Level 5",                          "—",                          "Level 4"
  )

tbl_improved_pts_caption <- "Country-year cases where including civil society restrictions improved PTS predictions"

if (!is_docx) {
  tbl_improved_pts %>% 
    kbl(align = "llcccc",
        booktabs = TRUE, linesep = "\\addlinespace",
        caption = tbl_improved_pts_caption) %>% 
    kable_styling(latex_options = c("hold_position")) %>% 
    row_spec(0, bold = TRUE) %>% 
    column_spec(1, width = "0.1\\\\textwidth") %>% 
    column_spec(2, width = "0.05\\\\textwidth") %>% 
    column_spec(3:6, width = "0.16\\\\textwidth")
} else {
  tbl_improved_pts %>% 
    hux() %>% 
    set_font_size(10) %>%
    set_all_padding(3) %>% 
    set_align(everywhere, 1:2, "left") %>% 
    set_align(everywhere, 3:6, "center") %>% 
    set_bold(1, everywhere) %>%
    set_bottom_border(1, everywhere) %>%
    set_width(1) %>% 
    set_wrap(TRUE) %>% 
    huxtable::set_caption(tbl_improved_pts_caption) %>% 
    huxtable::as_flextable()
}
```

\newpage

\blandscape

## Complete model results

\memSingleSmall

```{r tbl-models-pts}
model_names_tbl_pts <- c("Total barriers",
                         "Barriers to advocacy",
                         "Barriers to entry",
                         "Barriers to funding",
                         "Civil society repression")

names(models_paper_pts) <- model_names_tbl_pts

tbl_reg_pts_caption <- "Full results from ordered logistic regression models predicting political terror"

tbl_reg_pts <- modelsummary(models_paper_pts,
                            output = table_format,
                            statistic = "[{conf.low}, {conf.high}]",
                            coef_map = coef_list,
                            gof_omit = "ELPD",
                            escape = FALSE,
                            caption = tbl_reg_pts_caption,
                            notes = c(model_note),
                            longtable = TRUE)

if (!is_docx) {
  tbl_reg_pts %>% 
    kable_styling() %>% 
    add_header_above(header = c(" " = 1, "Outcome in t + 1" = 5))
    # kable_styling(font_size = ifelse(is_latex, 7, NA)) %>%
    # column_spec(1, width = "0.75in", latex_valign = "b") %>%
    # column_spec(2:6, width = "1.3in", latex_valign = "b")
    # column_spec(1, width = "0.10\\\\textwidth") %>%
    # column_spec(2:6, width = "0.14\\\\textwidth")
} else {
  tbl_reg_pts %>% 
    set_font_size(10) %>% 
    set_all_padding(1) %>% 
    set_align(everywhere, 2:6, "center") %>% 
    set_bold(1, everywhere) %>% 
    # set_col_width(1, 0.1) %>% 
    # set_col_width(2:6, 0.14) %>% 
    set_width(1) %>% 
    set_wrap(TRUE) %>% 
    huxtable::set_caption(tbl_reg_pts_caption) %>% 
    huxtable::as_flextable()
}
```

\clearpage

```{r tbl-models-lhr}
model_names_tbl_lhr <- c("Total barriers",
                         "Barriers to advocacy",
                         "Barriers to entry",
                         "Barriers to funding",
                         "Civil society repression")

names(models_paper_lhr) <- model_names_tbl_lhr

tbl_reg_lhr_caption <- "Full results from Gaussian regression models predicting latent human rights"

tbl_reg_lhr <- modelsummary(models_paper_lhr,
                            output = table_format,
                            statistic = "[{conf.low}, {conf.high}]",
                            coef_map = coef_list,
                            gof_omit = "ELPD",
                            escape = FALSE,
                            caption = tbl_reg_lhr_caption,
                            notes = c(model_note),
                            longtable = TRUE)

if (!is_docx) {
  tbl_reg_lhr %>% 
    kable_styling() %>% 
    add_header_above(header = c(" " = 1, "Outcome in t + 1" = 5))
    # kable_styling(font_size = ifelse(is_latex, 7.5, NA)) %>% 
    # column_spec(1, width = "0.10\\\\textwidth") %>% 
    # column_spec(2:6, width = "0.14\\\\textwidth")
} else {
  tbl_reg_lhr %>% 
    set_font_size(10) %>% 
    set_all_padding(1) %>% 
    set_align(everywhere, 2:6, "center") %>% 
    set_bold(1, everywhere) %>% 
    set_width(1) %>% 
    set_wrap(TRUE) %>% 
    huxtable::set_caption(tbl_reg_lhr_caption) %>% 
    huxtable::as_flextable()
}
```

\clearpage

\elandscape

\memSingle