### 2021-11-19 analysis/03_model-details.Rmd

---
title: "Model details and diagnostics"
author: "Suparna Chaudhry and Andrew Heiss"
date: "Last run: `r format(Sys.time(), '%F')`"
output: 
  html_document:
    code_folding: hide
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
library(knitr)
library(kableExtra)
knit_print.data.frame <- function(x, ...) {
  res <- paste(c('', '', kable_styling(kable(x, booktabs = TRUE))), collapse = '\n')
  asis_output(res)
}

registerS3method("knit_print", "data.frame", knit_print.data.frame)
registerS3method("knit_print", "grouped_df", knit_print.data.frame)

knitr::opts_chunk$set(fig.retina = 3,
                      tidy.opts = list(width.cutoff = 120),  # For code
                      options(width = 90),  # For output
                      fig.asp = 0.618, fig.width = 7, 
                      fig.align = "center", out.width = "85%")

options(dplyr.summarise.inform = FALSE,
        knitr.kable.NA = "")
```

```{r load-libraries-data, message=FALSE, warning=FALSE}
library(tidyverse)
library(targets)
library(broom)
library(broom.mixed)
library(tidybayes)
library(glue)
library(brms)
library(scales)
library(kableExtra)
library(modelsummary)
library(lubridate)
library(here)

# Generated via random.org
set.seed(9936)

# Load models
withr::with_dir(here::here(), {
  source(tar_read(plot_funs))

  # Load big list of models
  tar_load(model_df)

  # Load actual model objects
  # This works, but doesn't put the models in the dependency network for some reason
  # tar_load(model_df$model)

  # This also works but again doesn't register correctly as dependencies
  # tar_load(starts_with("m_pts"))

  # So we do it this manual way
  tar_load(c(m_pts_baseline, m_pts_total, m_pts_total_new, 
             m_pts_advocacy, m_pts_entry, m_pts_funding, 
             m_pts_v2csreprss, 
             m_lhr_baseline, m_lhr_total, m_lhr_total_new, 
             m_lhr_advocacy, m_lhr_entry, m_lhr_funding, 
             m_lhr_v2csreprss, 
             m_pts_baseline_train, m_pts_total_train, 
             m_pts_advocacy_train, m_pts_entry_train, m_pts_funding_train, 
             m_pts_v2csreprss_train,
             m_lhr_baseline_train, m_lhr_total_train, m_lhr_total_new_train, 
             m_lhr_advocacy_train, m_lhr_entry_train, m_lhr_funding_train, 
             m_lhr_v2csreprss_train))
})
```

```{r set-computer-details}
computer <- "Personal"

if (computer == "Work") {
  computer_details <- "2019 MacBook Pro with 16 cores and 32 GB of RAM, using Stan through brms through cmdstanr"
} else {
  computer_details <- "2021 MacBook Pro M1 Max with 10 cores and 32 GB of RAM, using Stan through brms through cmdstanr"
}
```

We ran these models on a `r computer_details`.

\

# Model run times

## Models for E~1~ (de jure civil society laws)

### Full data

```{r running-time-e1}
models_e1 <- model_df %>% 
  filter(!str_detect(model, "v2csreprss")) %>% 
  mutate(actual_model = model %>% map(~eval(rlang::sym(.)))) %>% 
  mutate(across(c(outcome_var, explan_var, family), ~fct_inorder(., ordered = TRUE)))

model_time_e1 <- models_e1 %>% 
  filter(training == "Full data") %>% 
  mutate(duration = map(actual_model, ~rstan::get_elapsed_time(.$fit)),
         duration = map(duration, ~rownames_to_column(as_tibble(.)))) %>% 
  select(-actual_model) %>% 
  unnest(duration) %>% 
  mutate(model = glue("<code>{model}</code>")) %>% 
  group_by(Model = model, Outcome = outcome_var, 
           `Main predictor` = explan_var, `Family` = family) %>% 
  summarize(`Total time (i.e. longest chain)` = as.duration(max(warmup + sample))) %>%
  ungroup() %>% 
  arrange(Outcome, `Main predictor`)

total_row_e1 <- tibble(Outcome = "Total", 
                       `Total time (i.e. longest chain)` = 
                         as.duration(sum(model_time_e1$`Total time (i.e. longest chain)`)))

model_time_e1 <- model_time_e1 %>% 
  bind_rows(total_row_e1)

model_time_e1 %>% 
  select(-Outcome) %>% 
  kbl(escape = FALSE) %>% 
  pack_rows(index = table(fct_inorder(model_time_e1$Outcome))) %>% 
  kable_styling()
```

### Training data

```{r running-time-e1-training}
model_time_e1 <- models_e1 %>% 
  filter(training == "Training") %>% 
  mutate(duration = map(actual_model, ~rstan::get_elapsed_time(.$fit)),
         duration = map(duration, ~rownames_to_column(as_tibble(.)))) %>% 
  select(-actual_model) %>% 
  unnest(duration) %>% 
  mutate(model = glue("<code>{model}</code>")) %>% 
  group_by(Model = model, Outcome = outcome_var, 
           `Main predictor` = explan_var, `Family` = family) %>% 
  summarize(`Total time (i.e. longest chain)` = as.duration(max(warmup + sample))) %>%
  ungroup() %>% 
  arrange(Outcome, `Main predictor`)

total_row_e1 <- tibble(Outcome = "Total", 
                       `Total time (i.e. longest chain)` = 
                         as.duration(sum(model_time_e1$`Total time (i.e. longest chain)`)))

model_time_e1 <- model_time_e1 %>% 
  bind_rows(total_row_e1)

model_time_e1 %>% 
  select(-Outcome) %>% 
  kbl(escape = FALSE) %>% 
  pack_rows(index = table(fct_inorder(model_time_e1$Outcome))) %>% 
  kable_styling()
```

## Models for E~2~ (de facto civil society environment)

### Full data

```{r running-time-e2}
models_e2 <- model_df %>% 
  filter(str_detect(model, "baseline") | str_detect(model, "v2csreprss")) %>% 
  mutate(actual_model = model %>% map(~eval(rlang::sym(.)))) %>% 
  mutate(across(c(outcome_var, explan_var, family), ~fct_inorder(., ordered = TRUE)))

model_time_e2 <- models_e2 %>% 
  filter(training == "Full data") %>% 
  mutate(duration = map(actual_model, ~rstan::get_elapsed_time(.$fit)),
         duration = map(duration, ~rownames_to_column(as_tibble(.)))) %>% 
  select(-actual_model) %>% 
  unnest(duration) %>% 
  mutate(model = glue("<code>{model}</code>")) %>% 
  group_by(Model = model, Outcome = outcome_var, 
           `Main predictor` = explan_var, `Family` = family) %>% 
  summarize(`Total time (i.e. longest chain)` = as.duration(max(warmup + sample))) %>%
  ungroup() %>% 
  arrange(Outcome, `Main predictor`)

total_row_e2 <- tibble(Outcome = "Total", 
                       `Total time (i.e. longest chain)` = 
                         as.duration(sum(model_time_e2$`Total time (i.e. longest chain)`)))

model_time_e2 <- model_time_e2 %>% 
  bind_rows(total_row_e2)

model_time_e2 %>% 
  select(-Outcome) %>% 
  kbl(escape = FALSE) %>% 
  pack_rows(index = table(fct_inorder(model_time_e2$Outcome))) %>% 
  kable_styling()
```

### Training data

```{r running-time-e2-training}
model_time_e2 <- models_e2 %>% 
  filter(training == "Training") %>% 
  mutate(duration = map(actual_model, ~rstan::get_elapsed_time(.$fit)),
         duration = map(duration, ~rownames_to_column(as_tibble(.)))) %>% 
  select(-actual_model) %>% 
  unnest(duration) %>% 
  mutate(model = glue("<code>{model}</code>")) %>% 
  group_by(Model = model, Outcome = outcome_var, 
           `Main predictor` = explan_var, `Family` = family) %>% 
  summarize(`Total time (i.e. longest chain)` = as.duration(max(warmup + sample))) %>%
  ungroup() %>% 
  arrange(Outcome, `Main predictor`)

total_row_e2 <- tibble(Outcome = "Total", 
                       `Total time (i.e. longest chain)` = 
                         as.duration(sum(model_time_e2$`Total time (i.e. longest chain)`)))

model_time_e2 <- model_time_e2 %>% 
  bind_rows(total_row_e2)

model_time_e2 %>% 
  select(-Outcome) %>% 
  kbl(escape = FALSE) %>% 
  pack_rows(index = table(fct_inorder(model_time_e2$Outcome))) %>% 
  kable_styling()
```

\

# Actual code

All the models are run with a **`targets`** pipeline with dataset-specific functions that live in these files:

- `R/models_pts.R` (for PTS-based models)
- `R/models_lhr.R` (for latent human rights-based models)

To keep the analysis relatively self contained here in the analysis notebook, and to make it so there's no need to hunt through the GitHub repository to find the actual code, here's that code:

#### `R/models_pts.R`

```{r, code=xfun::read_utf8(here::here("R", "models_pts.R")), eval=FALSE, class.source="fold-hide"}
```

#### `R/models_lhr.R`

```{r, code=xfun::read_utf8(here::here("R", "models_lhr.R")), eval=FALSE, class.source="fold-hide"}
```